---
title: "R vs Python"
subtitle: "Equivalent common data analysis functions in R and Python"
author: "Inmaculada Ruiz-Morales"
date: "`r Sys.Date()`"
output: html_document
theme: "cerulean" 
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE )
```


For this exercise, I am going to use a dataset containing information on community pharmacy activity in Scotland, covering February 2023 for all NHS health boards. Data was retrieved from Public Health Scotland website https://www.opendata.nhs.scot/dataset/prescriptions-in-the-community/resource/ad6d7f4e-08fd-40ed-ad49-bf450d386f39?inner_span=True


## To write Python in RMarkdown

First, you will need to install the reticulate package, after that, you can write python code using the {python} chunk header.*the header {python} will automatically use the reticulate package.*


## Importing packages 

```{r}
library(tidyverse)
```

```{python}
import numpy as np
import pandas as pd
```


## read csv files 


Be patient...it will take some time to load the data 

```{r}
# reading the data with R
r_path = "https://www.opendata.nhs.scot/dataset/84393984-14e9-4b0d-a797-b288db64d088/resource/ad6d7f4e-08fd-40ed-ad49-bf450d386f39/download/pitc202302.csv"
r_data <- read_csv(r_path)
```


```{python}
# reading the data with phyton 
py_path = "https://www.opendata.nhs.scot/dataset/84393984-14e9-4b0d-a797-b288db64d088/resource/ad6d7f4e-08fd-40ed-ad49-bf450d386f39/download/pitc202302.csv"
py_data = pd.read_csv(py_path)
```

## to see the first few rows and the last ones

```{r}
head(r_data)
```

```{r}
tail(r_data)
```

```{python}
py_data.head()
```

```{python}
py_data.tail()
```

## Find what type of object we have 

```{r}
class(r_data)
```

```{python}
type(py_data)
```

## Dimensions of the dataset.

```{r}
dim(r_data)
```

```{python}
py_data.shape
```


## Summary of the dataset.

```{r}
str(r_data)
```


```{python}
py_data.info()
```


## Summary of the statistical properties of the dataset.

```{r}
summary(r_data)
```

```{python}
py_data.describe()
```

## print column names

```{r}
colnames(r_data)
```

```{python}
py_data.columns
```

## to see unique categories in a variable

```{r}
unique(r_data$HBT)
```

```{python}
pd.unique(py_data["HBT"])
```

## Counts values for items dispensed

```{r}
r_data %>% 
  count(r_data$BNFItemDescription)
```

Python gives you the values sorted in descending order. 

```{python}
py_data['BNFItemDescription'].value_counts()
```

Lets see the 40 most frequently dispensed items in February 2023

```{python}
Pmost_dispensed =  py_data['BNFItemDescription'].value_counts()
Pmost_dispensed.head(40)
```


Same result with R code. *(notice that R is also counting the NA)*  

```{r}
Rmost_dispensed <- r_data %>%
  count(r_data$BNFItemDescription) %>%
  arrange(desc(n))
head(Rmost_dispensed, 40) %>%
  kable()
```


```{r}
Rmost_dispensed <- r_data %>%
  count(r_data$BNFItemDescription) %>%
  arrange(desc(n)) %>%
  setNames(c("BNF Item Description", "Frecuency"))

head(Rmost_dispensed, 40) %>%
  kable()

```




>www.inmaruiz.com





